pseudobulk_CMP

### 2021-09-13 pseudobulk_vs_bulk/pseudobulk_CMP.Rmd
appended TPM vs CPM correlation tutorial

## Set global variables
```{r}
projectName <- "pseudobulk_CMP"
```

@@

>>>>

## Set global variables
```{r}
projectName <- "pseudobulk_CMP"
comparison_pop <- "CMP"
```

@@

# Single cell/Bulk RNA-Seq concordance tutorial
Following a new (tutorial) [https://jdblischak.github.io/singlecell-qtl/sc-vs-bulk.html].

Calculate cpm from single cell umi matrix
```{r}
sc.mtx <- data.frame(rowSums(seurat.object@assays$RNA@counts))
sc.mtx <- sc.mtx/colSums(sc.mtx)*1000000
```
Tutorial states that only quantity directly cmoprable between bulk and scRNA-Seq is relative abundance (cites Pachter 2001).

read cmp bulk TPM
```{r}
# generate table with individual TPMs
bulk.path <- "../../bRNASeq/TotalSeq/"
file_list <- list.files(path = bulk.path, pattern = comparison_pop)
bulk.pop <- read.table(file = paste0(bulk.path, file_list[1]), header = TRUE)[1]
for(file in file_list){
	print(file)
	bulk.pop <- merge(bulk.pop, read.table(file = paste0(bulk.path, file), header = TRUE, sep = "\t")[, c(1,5)], by = "gene_id")
}

#calculate average tpm
bulk.pop[comparison_pop] <- rowMeans(x = bulk.pop[, c(2, 3)])

# clean and check df
bulk.pop <- bulk.pop[, -c(2, 3)]
head(bulk.pop)
```
```{r}
#calc log_rho's
bulk.rho <- log((bulk.pop[comparison_pop] + 1), 2) - log(colSums((bulk.pop[comparison_pop])+ 1), 2)
rownames(bulk.rho) <- bulk.pop$gene_id
colnames(bulk.rho) <- "bulk"
sc.rho <- log((sc.mtx + 1), 2) - log(colSums((sc.mtx+1)), 2)
colnames(sc.rho) <- "sc"
```

check that all values are numeric
```{r}
print(is.numeric(bulk.rho[,1]))
print(is.numeric(sc.rho[,1]))
```
## create concordance df
### set NA = 0
```{r}
concordance.NAasZero <- merge(bulk.rho, sc.rho, by = "row.names", all = TRUE)
concordance.NAasZero[is.na(concordance.NAasZero)] <- 0
rownames(concordance.NAasZero) <- concordance.NAasZero$Row.names
concordance.NAasZero <- concordance.NAasZero[, c(2, 3)]
```
#### plot data
```{r}
ggplot(data = concordance.NAasZero, aes(x = bulk, y = sc)) + 
	geom_point()  +
	ggtitle(paste0("concordance in ", comparison_pop))
```
```{r}
cor(concordance.NAasZero, method = "spearman")
```

### drop NA
```{r}
concordance.df <- merge(bulk.rho, sc.rho, by = "row.names", all = TRUE)
concordance.df <- na.omit(concordance.df)
rownames(concordance.df) <- concordance.df$Row.names
concordance.df <- concordance.df[, c(2, 3)]
```
```{r}
spearman.coeff <- cor(concordance.df, method = "spearman")[1, 2]
spearman.coeff
```
#### plot data
```{r}
ggplot(data = concordance.df, aes(x = bulk, y = sc)) + 
	geom_point()  +
	ggtitle(paste0("concordance in ", comparison_pop, " spearman coeff = ", round(spearman.coeff, digits = 3)))
```
